---
title: "Correlation testing"
output: html_notebook
---

This notebook is for testing ONLY. It focuses on two datasets, to test how to do the correlation analysis.



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(dplyr)
library(Taba)
library(XICOR)
```



# Load and reformat data
Load Seurat data

```{r}
lizard_neurons_v3 = readRDS("data/expression/lizard_neurons_v3.RDS")
turtle_neurons_v3 = readRDS("data/expression/turtle_neurons_v3.RDS")
```

Load necessary eggNOG data

```{r}
annot_list = readRDS(file = "data/eggNOG/annotation_list.RDS")
```

Reformat matrices to have only one2one orthologs that are shared

```{r}
exp_tur = turtle_neurons_v3@assays$RNA@counts
exp_liz = lizard_neurons_v3@assays$RNA@counts

turtle_egg = annot_list$Cpicta[,c(2,5)]
turtle_egg = turtle_egg[turtle_egg$Preferred_name!="-" & !is.na(turtle_egg$Preferred_name),]
turtle_egg$Preferred_name = toupper(turtle_egg$Preferred_name)
turtle_egg = unique(turtle_egg)
turtle_egg = turtle_egg[!(turtle_egg$gene %in% turtle_egg$gene[duplicated(turtle_egg$gene)]),]
turtle_egg = turtle_egg[!(turtle_egg$Preferred_name %in% turtle_egg$Preferred_name[duplicated(turtle_egg$Preferred_name)]),]
lizard_egg = annot_list$Pvitticeps[,c(2,5)]
lizard_egg = lizard_egg[lizard_egg$Preferred_name!="-" & !is.na(lizard_egg$Preferred_name),]
lizard_egg$Preferred_name = toupper(lizard_egg$Preferred_name)
lizard_egg = unique(lizard_egg)
lizard_egg = lizard_egg[!(lizard_egg$gene %in% lizard_egg$gene[duplicated(lizard_egg$gene)]),]
lizard_egg = lizard_egg[!(lizard_egg$Preferred_name %in% lizard_egg$Preferred_name[duplicated(lizard_egg$Preferred_name)]),]

liztur_egg = merge(lizard_egg, turtle_egg, by = 2)
colnames(liztur_egg) = c("Preferred_name", "gene_lizard", "gene_turtle")
liztur_egg = liztur_egg[liztur_egg$gene_lizard %in% rownames(exp_liz) &
                          liztur_egg$gene_turtle %in% rownames(exp_tur),]

exp_tur = exp_tur[rownames(exp_tur) %in% liztur_egg$gene_turtle,]
rownames(exp_tur) = liztur_egg$Preferred_name[match(rownames(exp_tur), liztur_egg$gene_turtle)]
exp_liz = exp_liz[rownames(exp_liz) %in% liztur_egg$gene_lizard,]
rownames(exp_liz) = liztur_egg$Preferred_name[match(rownames(exp_liz), liztur_egg$gene_lizard)]
exp_liz = exp_liz[rownames(exp_tur),]
dim(exp_tur)
dim(exp_liz)
```

Renormalise

```{r}
turtle_orth = CreateSeuratObject(exp_tur, meta.data = turtle_neurons_v3@meta.data)
Idents(turtle_orth) = turtle_orth@meta.data$clusters
turtle_orth = suppressWarnings(SCTransform(turtle_orth, do.correct.umi = T, verbose = F, seed.use = 1,
                                           vars.to.regress = "nCount_RNA", variable.features.rv.th = 1, 
                                           return.only.var.genes = F, variable.features.n = NULL))
lizard_orth = CreateSeuratObject(exp_liz, meta.data = lizard_neurons_v3@meta.data)
Idents(lizard_orth) = lizard_orth@meta.data$clusters
lizard_orth = suppressWarnings(SCTransform(lizard_orth, do.correct.umi = T, verbose = F, seed.use = 1,
                                           vars.to.regress = "nCount_RNA", variable.features.rv.th = 1, 
                                           return.only.var.genes = F, variable.features.n = NULL))

turtle_avg = AverageExpression(turtle_orth, assays = "SCT")$SCT
lizard_avg = AverageExpression(lizard_orth, assays = "SCT")$SCT
# get clusters of interest (similar to paper)
tur_cl = unique(turtle_orth@meta.data$clusters[turtle_orth@meta.data$pallial.area %in% c("MC", "DMC")])
liz_cl = unique(lizard_orth@meta.data$clusters[lizard_orth@meta.data$pallial.area %in% c("MC", "DMC")])
```

Get marker genes

```{r}
tur_mk = FindAllMarkers(turtle_orth, logfc.threshold = 0.2, pseudocount.use = 0.1)
liz_mk = FindAllMarkers(lizard_orth, logfc.threshold = 0.2, pseudocount.use = 0.1)

saveRDS(tur_mk, file = "results/correlation_testing/turtle_markers.RDS")
saveRDS(liz_mk, file = "results/correlation_testing/lizard_markers.RDS")

tur_mk = FindAllMarkers(turtle_orth[,turtle_orth$clusters %in% tur_cl], 
                        logfc.threshold = 0.2, pseudocount.use = 0.1)
liz_mk = FindAllMarkers(lizard_orth[,lizard_orth$clusters %in% liz_cl], 
                        logfc.threshold = 0.2, pseudocount.use = 0.1)

saveRDS(tur_mk, file = "results/correlation_testing/turtle_subset_markers.RDS")
saveRDS(liz_mk, file = "results/correlation_testing/lizard_subset_markers.RDS")
```

Select genes to use for correlations, normalise data

```{r}
# get top markers
tur_mk_use = tur_mk[tur_mk$p_val_adj<=0.05 & tur_mk$avg_logFC>0.2,]
liz_mk_use = liz_mk[liz_mk$p_val_adj<=0.05 & liz_mk$avg_logFC>0.2,]
# top genes per cluster
tur_mk_top = unique(unlist((tur_mk_use %>% arrange(desc(avg_logFC)) %>% group_by(cluster) %>% slice(1:500000))[,"gene"]))
liz_mk_top = unique(unlist((liz_mk_use %>% arrange(desc(avg_logFC)) %>% group_by(cluster) %>% slice(1:500000))[,"gene"]))

common_genes = intersect(tur_mk_top, liz_mk_top)
common_genes = common_genes[common_genes %in% intersect(rownames(turtle_avg), rownames(lizard_avg))]
turtle_avg_sub = turtle_avg[common_genes,]
turtle_avg_sub = t(apply(turtle_avg_sub, 1, function(x) x/mean(x)))
lizard_avg_sub = lizard_avg[common_genes,]
lizard_avg_sub = t(apply(lizard_avg_sub, 1, function(x) x/mean(x)))

dim(lizard_avg_sub)
```

Calculate correlations

```{r, include = FALSE}
list_cor = list("Spearman" = cor(turtle_avg_sub, lizard_avg_sub, method = "sp"),
                "Pearson" = cor(turtle_avg_sub, lizard_avg_sub, method = "pe"))

xicor_mat = matrix(0, ncol = ncol(lizard_avg_sub), nrow = ncol(turtle_avg_sub))
xicor_rev = matrix(0, ncol = ncol(turtle_avg_sub), nrow = ncol(lizard_avg_sub))
rownames(xicor_mat) = colnames(xicor_rev) = colnames(turtle_avg_sub)
colnames(xicor_mat) = rownames(xicor_rev) = colnames(lizard_avg_sub)
for(i in 1:nrow(xicor_mat)){
  for(j in 1:ncol(xicor_mat)){
    xicor_mat[i,j] = xicor(turtle_avg_sub[,i], lizard_avg_sub[,j])
    xicor_rev[j,i] = xicor(lizard_avg_sub[,j], turtle_avg_sub[,i])
  }
}
list_cor$xicor_tl = xicor_mat
list_cor$xicor_lt = t(xicor_rev)

taba_mat = matrix(0, ncol = ncol(lizard_avg_sub), nrow = ncol(turtle_avg_sub))
tabarank_mat = matrix(0, ncol = ncol(lizard_avg_sub), nrow = ncol(turtle_avg_sub))
tabwil_mat = matrix(0, ncol = ncol(lizard_avg_sub), nrow = ncol(turtle_avg_sub))
tabwilrank_mat = matrix(0, ncol = ncol(lizard_avg_sub), nrow = ncol(turtle_avg_sub))
rownames(taba_mat) = rownames(tabarank_mat) = rownames(tabwil_mat) = rownames(tabwilrank_mat) = colnames(turtle_avg_sub)
colnames(taba_mat) = colnames(tabarank_mat) = colnames(tabwil_mat) = colnames(tabwilrank_mat) = colnames(lizard_avg_sub)
for(i in 1:nrow(taba_mat)){
  for(j in 1:ncol(taba_mat)){
    taba_mat[i,j] = taba(turtle_avg_sub[,i], lizard_avg_sub[,j], method = "taba")
    tabarank_mat[i,j] = taba(turtle_avg_sub[,i], lizard_avg_sub[,j], method = "tabarank")
    tabwil_mat[i,j] = taba(turtle_avg_sub[,i], lizard_avg_sub[,j], method = "tabwil")
    tabwilrank_mat[i,j] = taba(turtle_avg_sub[,i], lizard_avg_sub[,j], method = "tabwilrank")
  }
}
list_cor$taba = taba_mat
list_cor$tabarank = tabarank_mat
list_cor$tabwil = tabwil_mat
list_cor$tabwilrank = tabwilrank_mat
```

Plot correlation matrices

```{r}
turtle_annot = unique(turtle_orth@meta.data[,c("pallial.area","clusters")])
turtle_annot = data.frame(row.names = turtle_annot$clusters, "pallial.area" = turtle_annot$pallial.area)
lizard_annot = unique(lizard_orth@meta.data[,c("pallial.area","clusters")])
lizard_annot = data.frame(row.names = lizard_annot$clusters, "pallial.area" = lizard_annot$pallial.area)

cols = colorRampPalette(c(rev(RColorBrewer::brewer.pal(9, "Blues")),
                          RColorBrewer::brewer.pal(9, "Reds")))(100)
for(n in names(list_cor)){
  pltmat = t(list_cor[[n]][tur_cl[order(tur_cl)][c(1:7,9,10,8)],
                                     liz_cl[order(liz_cl)][c(1:5,8,9,6,7)]])
  br = c(seq(min(pltmat), 0, length.out = 51), seq(0, max(pltmat), length.out = 51)[-1])
  #br = seq(min(pltmat), max(pltmat), length.out = 51)
  pheatmap::pheatmap(pltmat, breaks = br, color = cols,
                     clustering_method = "ward.D2", main = n, 
                     annotation_row = lizard_annot, annotation_col = turtle_annot)
  
  pheatmap::pheatmap(pltmat, breaks = br, color = cols,
                     cluster_rows = F, cluster_cols = F, main = n, gaps_row = c(5,7), gaps_col = c(4,6),
                     annotation_row = lizard_annot, annotation_col = turtle_annot)
}
```

Calculate correlations for genes

```{r, include = FALSE}
turtle_avg_sub = turtle_avg[common_genes,]
turtle_avg_sub = apply(turtle_avg_sub, 1, function(x) x/mean(x))

list_cor = list("Spearman" = cor(turtle_avg_sub, method = "sp"),
                "Pearson" = cor(turtle_avg_sub, method = "pe"),
                "xicor" = xicor(turtle_avg_sub))

taba_mat = matrix(0, ncol = ncol(turtle_avg_sub), nrow = ncol(turtle_avg_sub))
tabarank_mat = matrix(0, ncol = ncol(turtle_avg_sub), nrow = ncol(turtle_avg_sub))
tabwil_mat = matrix(0, ncol = ncol(turtle_avg_sub), nrow = ncol(turtle_avg_sub))
tabwilrank_mat = matrix(0, ncol = ncol(turtle_avg_sub), nrow = ncol(turtle_avg_sub))
rownames(taba_mat) = rownames(tabarank_mat) = rownames(tabwil_mat) = rownames(tabwilrank_mat) = colnames(turtle_avg_sub)
colnames(taba_mat) = colnames(tabarank_mat) = colnames(tabwil_mat) = colnames(tabwilrank_mat) = colnames(turtle_avg_sub)
for(i in 1:nrow(taba_mat)){
  for(j in 1:ncol(taba_mat)){
    taba_mat[i,j] = taba(turtle_avg_sub[,i], turtle_avg_sub[,j], method = "taba")
    tabarank_mat[i,j] = taba(turtle_avg_sub[,i], turtle_avg_sub[,j], method = "tabarank")
    tabwil_mat[i,j] = taba(turtle_avg_sub[,i], turtle_avg_sub[,j], method = "tabwil")
    tabwilrank_mat[i,j] = taba(turtle_avg_sub[,i], turtle_avg_sub[,j], method = "tabwilrank")
  }
}
list_cor$taba = taba_mat
list_cor$tabarank = tabarank_mat
list_cor$tabwil = tabwil_mat
list_cor$tabwilrank = tabwilrank_mat
```

Plot correlation matrices

```{r, fig.height=5, fig.width=5.5}
cols = colorRampPalette(c(rev(RColorBrewer::brewer.pal(9, "Blues")),
                          RColorBrewer::brewer.pal(9, "Reds")))(100)
for(n in names(list_cor)){
  pltmat = list_cor[[n]]
  if(!any(is.na(pltmat))){
    diag(pltmat) = 0
    br = c(seq(min(pltmat), 0, length.out = 51), seq(0, max(pltmat), length.out = 51)[-1])
    diag(pltmat) = NA
    pheatmap::pheatmap(pltmat, breaks = br, color = cols,
                       clustering_method = "ward.D2", main = n)
  }
}
```

Look at high subset in xicor

```{r, fig.height=6, fig.width=8}
pltmat = list_cor$xicor
diag(pltmat) = NA
hcl = hclust(dist(t(pltmat)), method = "ward.D2")
colm = cutree(hcl, 2)
hcl = hclust(dist(pltmat), method = "ward.D2")
rows = cutree(hcl, 2)
both = cbind(rows, colm)
both = both[both[,1]==2 | both[,2]==2,]
sub_pltmat = list_cor$xicor[names(rows)[rows==2], names(colm)[colm==2]]
for(i in unique(rownames(sub_pltmat))){
  for(j in unique(colnames(sub_pltmat))){
    if(i==j){ sub_pltmat[i,j] = 0 }
  }
}
br = c(seq(min(sub_pltmat), 0, length.out = 51), seq(0, max(sub_pltmat), length.out = 51)[-1])
for(i in unique(rownames(sub_pltmat))){
  for(j in unique(colnames(sub_pltmat))){
    if(i==j){ sub_pltmat[i,j] = NA }
  }
}
pheatmap::pheatmap(sub_pltmat, breaks = br, color = cols,
                   clustering_method = "ward.D2", fontsize_row = 7.6, fontsize_col = 7.6)
```

Check genes with greates differences

```{r}
xxx = reshape2::melt(list_cor$xicor)

diffv = c()
for(i in unique(xxx$Var1)){
  for(j in unique(xxx$Var2)){
    diffv = c(diffv, abs(xxx[xxx$Var1==i & xxx$Var2==j,3]-xxx[xxx$Var2==i & xxx$Var1==j,3]))
  }
}
xxx$diffv = diffv

View(xxx[xxx$Var1 %in% rownames(both) & xxx$Var2 %in% rownames(both),])
```




