---
title: "Cross species correlation - pairwise genes (updated orthology)"
output: html_notebook
---

Notebook for cell type identity correlations using pairwise genes. Subsetting will be needed since for some of the species we only used the first cell type name (which is what we're more interested in).



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(uwot)
```



# Load data
Load average expression data

```{r}
means_list = list(
  zebrafish = readRDS("results/pairwiseDE/drerio_avgExp.RDS")$subclasses,
  axolotl = readRDS("results/pairwiseDE/axolotlNuc_avgExp.RDS")$cellclusters,
  paintedturtle = readRDS("results/pairwiseDE/turtle_avgExp.RDS")$subclasses,
  lizard = readRDS("results/pairwiseDE/lizard_avgExp.RDS")$subclasses,
  zebrafinch = readRDS("results/pairwiseDE/zfinch_avgExp.RDS")$subclasses,
  bengalesefinch = readRDS("results/pairwiseDE/bfinch_avgExp.RDS")$subclasses,
  mouse = readRDS("results/pairwiseDE/mouse_avgExp.RDS")$subclasses,
  human = readRDS("results/pairwiseDE/human10x_avgExp.RDS")$subclasses
)
```

Load pairwise DE results

```{r}
pwde_list = list(
  dr = readRDS("results/pairwiseDE/drerio_markers.RDS")$subclasses,
  ax = readRDS("results/pairwiseDE/axolotlNuc_markers.RDS")$cellclusters,
  tu = readRDS("results/pairwiseDE/turtle_markers.RDS")$subclasses,
  pv = readRDS("results/pairwiseDE/lizard_markers.RDS")$subclasses,
  zf = readRDS("results/pairwiseDE/zfinch_markers.RDS")$subclasses,
  bf = readRDS("results/pairwiseDE/bfinch_markers.RDS")$subclasses,
  mm = readRDS("results/pairwiseDE/mouse_markers.RDS")$subclasses,
  hu = readRDS("results/pairwiseDE/human10x_markers.RDS")$subclasses
)
names(pwde_list) = names(means_list)
```

## Prepare orthologs matching
Load necessary orthology data

```{r}
ort_alt_ens_list = readRDS(file = "data/eggNOG/ort_alt_ens_list.RDS")
ort_alt_egg_list = readRDS(file = "data/eggNOG/ort_alt_egg_list.RDS")

ort_list_use = c(ort_alt_ens_list, 
                 ort_alt_egg_list[which(grepl("axolotl", names(ort_alt_egg_list)) |
                                          grepl("lizard", names(ort_alt_egg_list)))])
bfmats = ort_alt_egg_list[which(grepl("zebrafinch", names(ort_alt_egg_list)))]
names(bfmats) = gsub("zebrafinch", "bengalesefinch", names(bfmats))
ort_list_use = c(ort_list_use, bfmats)
```



# Comparisons
Function to match cells from 2 species and compare

```{r}
corrCellTypesPW = function(means_list, pwde_list, ort_l, ort_scope = "ortholog_one2one", 
                           sp_pair = c("human", "mouse"), 
                           ct_sets = NULL){
  # which cells to use
  if(is.null(ct_sets)){
    ct_sets = list(colnames(means_list[[sp_pair[1]]]), colnames(means_list[[sp_pair[2]]]))
    names(ct_sets) = sp_pair
  }
  
  # choose species orthologs and scope
  ort_tab = ort_l[[paste0(sp_pair, collapse = ".vs.")]]
  if(!is.null(ort_scope)){
    ort_tab = ort_tab[ort_tab$homology.type==ort_scope,]
  }
  ort_tab = unique(ort_tab[ort_tab[,2] %in% rownames(means_list[[sp_pair[1]]]) &
                             ort_tab[,4] %in% rownames(means_list[[sp_pair[2]]]),c(2,4)])
  
  # get DE genes for all cell types being considered
  de_genes_pw = list()
  for(sp in sp_pair){
    comp = names(pwde_list[[sp]])
    # select those comparing to each other
    compuse = rowSums(Reduce(cbind, lapply(ct_sets[[sp]], function(x) grepl(x, comp, fixed = T))))>1
    de_genes_pw[[sp]] = unique(unlist(lapply(pwde_list[[sp]][compuse], rownames)))
  }
  # using marker intersection
  ort_tab = ort_tab[ort_tab[,1] %in% de_genes_pw[[1]] & ort_tab[,2] %in% de_genes_pw[[2]],]
  
  # normalise means
  m_list_sub = lapply(means_list[which(names(means_list) %in% sp_pair)], 
                      function(x) t(apply(x, 1, function(y) y/mean(y))))
  # get correlations, determine max per row and column
  cort = psych::corr.test(m_list_sub[[sp_pair[1]]][ort_tab[,1],ct_sets[[1]]], 
                          m_list_sub[[sp_pair[2]]][ort_tab[,2],ct_sets[[2]]], method = "sp", 
                          adjust = "fdr", alpha = 0.01, ci = F)
  cort$maxrow = apply(cort$r, 1, which.max)
  cort$maxcol = apply(cort$r, 2, which.max)
  cort[[sp_pair[1]]] = m_list_sub[[sp_pair[1]]][ort_tab[,1],ct_sets[[1]]]
  cort[[sp_pair[2]]] = m_list_sub[[sp_pair[2]]][ort_tab[,2],ct_sets[[2]]]
  return(cort)
}
```

Plot correlations

```{r}
plotCorr = function(cort, sp1 = "sp1", sp2 = "sp2"){
  # cluster and order labels
  hcr = hclust(dist(cort$r), method = "ward.D2")
  hcc = hclust(dist(t(cort$r)), method = "ward.D2")
  hcr = hcr$labels[hcr$order]
  hcc = hcc$labels[hcc$order]
  
  # reshaping the correlations
  plot_df = reshape2::melt(cort$r)
  plot_df$Var1 = factor(plot_df$Var1, levels = rev(hcr))
  plot_df$Var2 = factor(plot_df$Var2, levels = hcc)
  
  # add pvalue and max cor infor
  plot_df$padj = -log10(reshape2::melt(cort$p.adj+min(cort$p.adj[cort$p.adj>0])/10)$value)
  plot_df$rowmax = apply(Reduce(cbind, lapply(names(cort$maxrow), 
                                              function(n) plot_df$Var1==n &
                                                plot_df$Var2==colnames(cort$r)[cort$maxrow[n]])), 
                         1, any)
  plot_df$colmax = apply(Reduce(cbind, lapply(names(cort$maxcol), 
                                              function(n) plot_df$Var2==n &
                                                plot_df$Var1==rownames(cort$r)[cort$maxcol[n]])), 
                         1, any)
  plot_df$markcol = plot_df$value>quantile(plot_df$value, 0.98)
  
  # getting a colourscale where 0 is white in the middle, and intensity leveled by max(abs(value))
  cols = colorRampPalette(c(rev(RColorBrewer::brewer.pal(9, "Blues")),
                            RColorBrewer::brewer.pal(9, "Reds")))(101)
  br = seq(-max(abs(cort$r)), max(abs(cort$r)), length.out = 101)
  cols = cols[!(br>max(cort$r) | br<min(cort$r))]
  
  corplot = ggplot()+
    geom_point(data = plot_df, mapping = aes(x = Var2, y = Var1, fill = value, size = padj), 
               shape = 21)+
    geom_point(data = plot_df[plot_df$rowmax,], mapping = aes(x = Var2, y = Var1, size = padj), 
               shape = "—", show.legend = F, colour = "grey10")+
    geom_point(data = plot_df[plot_df$colmax,], mapping = aes(x = Var2, y = Var1, size = padj), 
               shape = "|", show.legend = F, colour = "grey10")+
    scale_x_discrete(expand = c(0,0.7))+
    scale_y_discrete(expand = c(0,0.7))+
    scale_fill_gradientn(breaks = signif(c(min(cort$r)+0.005, 0, max(cort$r)-0.005),2), 
                         values = scales::rescale(c(min(br), 0, max(br))),
                         colours = cols)+
    labs(x = sp2, y = sp1, fill = "Spearman's\nrho", size = "-log10\nadj. p-value")+
    theme_classic()+
    theme(axis.title = element_text(colour = "black", face = "bold"),
          axis.text = element_text(colour = "black"),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 8))
  
  return(corplot)
}
```

Usage examples

```{r}
# using all cells
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "lizard"), ct_sets = NULL)
plotCorr(ccc, sp1 = "axolotl", sp2 = "lizard")

# using all cells
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "paintedturtle"), ct_sets = NULL)
plotCorr(ccc, sp1 = "axolotl", sp2 = "paintedturtle")

# using all cells
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("lizard", "paintedturtle"), ct_sets = NULL)
plotCorr(ccc, sp1 = "lizard", sp2 = "paintedturtle")



# using a cell subset (can be seen by looking at colnames of mean_list elements)
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "lizard"), 
                      ct_sets = list("axolotl" = colnames(means_list$axolotl)[grepl("glut_", colnames(means_list$axolotl))],
                                     "lizard" = c(paste0("pvExc", 1:4), "pvNPC")))
plotCorr(ccc, sp1 = "axolotl", sp2 = "lizard")
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "lizard"), 
                      ct_sets = list("axolotl" = colnames(means_list$axolotl)[grepl("GABA_", colnames(means_list$axolotl))],
                                     "lizard" = c(paste0("pvInh", 1:2), "pvNPC")))
plotCorr(ccc, sp1 = "axolotl", sp2 = "lizard")
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "lizard"), 
                      ct_sets = list("axolotl" = colnames(means_list$axolotl)[grepl("neuro_", colnames(means_list$axolotl))],
                                     "lizard" = c(paste0("pvInh", 1:2), paste0("pvExc", 1:4), "pvNPC")))
plotCorr(ccc, sp1 = "axolotl", sp2 = "lizard")

# using a cell subset (can be seen by looking at colnames of mean_list elements)
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("axolotl", "paintedturtle"), 
                      ct_sets = list("axolotl" = paste0("Glut_", 0:9),
                                     "paintedturtle" = c(paste0("tsExc", 1:5), "tsNPCs")))
plotCorr(ccc, sp1 = "axolotl", sp2 = "paintedturtle")

# using a cell subset (can be seen by looking at colnames of mean_list elements)
ccc = corrCellTypesPW(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      sp_pair = c("lizard", "paintedturtle"), 
                      ct_sets = list("lizard" = c(paste0("pvExc", 1:4), "pvNPC"),
                                     "paintedturtle" = c(paste0("tsExc", 1:5), "tsNPCs")))
plotCorr(ccc, sp1 = "lizard", sp2 = "paintedturtle")
```

Calculate and plot UMAP of cells

```{r}
umapCellTypes = function(means_list, pwde_list, ort_l, ort_scope = "ortholog_one2one", 
                         ct_sets = list("human" = "all", "mouse" = "all"), seed = 1){ 
  # species are detected from ct_sets names
  # which cells to use
  for(sp in names(ct_sets)){
    if(ct_sets[[sp]][1]=="all"){ 
      ct_sets[[sp]] = colnames(means_list[[sp]])
    }
  }
  
  # get orthologs for all species pairs
  sp_p_mat = combn(names(ct_sets), 2)
  for(i in 1:ncol(sp_p_mat)){
    sp_pair = sp_p_mat[,i]
    ort_tab = ort_l[[paste0(sp_pair, collapse = ".vs.")]]
    if(!is.null(ort_scope)){
      ort_tab = ort_tab[ort_tab$homology.type==ort_scope,]
    }
    # orthologs in the means matrix
    ort_tab = unique(ort_tab[ort_tab[,2] %in% rownames(means_list[[sp_pair[1]]]) &
                               ort_tab[,4] %in% rownames(means_list[[sp_pair[2]]]),c(2,4)])
    colnames(ort_tab) = gsub("genename", "", colnames(ort_tab))
    if(i==1){
      ort_all = ort_tab
    } else{
      ort_all = unique(merge(ort_all, ort_tab, by = intersect(colnames(ort_all), colnames(ort_tab))))
    }
  }
  
  # get DE genes for all cell types being considered
  de_genes_pw = list()
  for(sp in names(ct_sets)){
    comp = names(pwde_list[[sp]])
    # select those comparing to each other
    compuse = rowSums(Reduce(cbind, lapply(ct_sets[[sp]], function(x) grepl(x, comp, fixed = T))))>1
    de_genes_pw[[sp]] = unique(unlist(lapply(pwde_list[[sp]][compuse], rownames)))
  }
  keep = apply(data.frame(lapply(colnames(ort_all), function(x) ort_all[,x] %in% de_genes_pw[[x]])), 
               1, all)
  ort_all = ort_all[keep,]
  
  # normalise means
  m_list_sub = lapply(means_list[which(names(means_list) %in% names(ct_sets))], 
                      function(x) t(apply(x, 1, function(y) y/mean(y))))
  
  avg_exp_all = t(Reduce(cbind, lapply(names(m_list_sub), 
                                       function(x) m_list_sub[[x]][ort_all[,x],ct_sets[[x]]])))
  
  # define metadata
  meta = data.frame("cell_types" = unlist(lapply(m_list_sub, colnames)),
                    "species" = unlist(lapply(names(m_list_sub), 
                                              function(x) rep(x, ncol(m_list_sub[[x]])))))
  rownames(meta) = paste0(meta$species, "_", meta$cell_types)
  meta = meta[match(rownames(avg_exp_all), meta$cell_types),]
  rownames(avg_exp_all) = rownames(meta)
  
  # calculate UMAP
  set.seed(seed)
  l = uwot::umap(avg_exp_all, metric = "cosine", ret_nn = T, n_epochs = 1000)
  l_sp = data.frame(l$embedding)
  rownames(l_sp) = rownames(avg_exp_all)

  plot_df = cbind(l_sp, meta[rownames(l_sp),])
  
  return(plot_df)
}
```


```{r}
pltdf = umapCellTypes(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      ort_scope = NULL, ct_sets = list("lizard" = c(paste0("pvExc", 1:4), "pvNPC"),
                                                       "paintedturtle" = c(paste0("tsExc", 1:5),
                                                                           "tsNPCs"),
                                                       "axolotl" = paste0("Glut_", 0:9)))

ggplot(pltdf, aes(x = X1, y = X2, colour = species))+
  geom_point()+
  ggrepel::geom_text_repel(data = pltdf, mapping = aes(x = X1, y = X2, label = cell_types),
                           max.overlaps = 100)+
  theme_classic()+
  theme(aspect.ratio = 1)

pltdf = umapCellTypes(means_list = means_list, pwde_list = pwde_list, ort_l = ort_list_use,
                      ort_scope = NULL, ct_sets = list("lizard" = "all",
                                                       "paintedturtle" = "all",
                                                       "axolotl" = "all"))

ggplot(pltdf, aes(x = X1, y = X2, colour = species))+
  geom_point()+
  ggrepel::geom_text_repel(data = pltdf, mapping = aes(x = X1, y = X2, label = cell_types),
                           max.overlaps = 100)+
  theme_classic()+
  theme(aspect.ratio = 1)
```


UMAP from mean expression

```{r}
for(n in names(avg_exp_sub)){
  sp = strsplit(n, "_")[[1]][2]
  colnames(avg_exp_sub[[n]]) = paste0(sp, "..", colnames(avg_exp_sub[[n]]))
}
avg_exp_all = t(Reduce(cbind, avg_exp_sub))
set.seed(2954)
l = uwot::umap(avg_exp_all, metric = "cosine", ret_nn = T, n_epochs = 1000)
l_sp = data.frame(l$embedding)
rownames(l_sp) = rownames(avg_exp_all)
l_sp$cl = rownames(avg_exp_all)
l_sp$sp = unlist(lapply(strsplit(l_sp$cl, "..", fixed = T), function(x) x[1]))
l_sp$cl = unlist(lapply(strsplit(l_sp$cl, "..", fixed = T), function(x) x[2]))

df_names = l_sp[l_sp$sp=="axo" | (l_sp$X2<(-4) & l_sp$sp=="mou") | (l_sp$X2>0.5 & l_sp$sp=="mou"),]
ggplot(l_sp, aes(x = X1, y = X2, colour = sp))+
  geom_point()+
  ggrepel::geom_text_repel(data = df_names, mapping = aes(x = X1, y = X2, label = cl), max.overlaps = 100)+
  theme_classic()
```


