---
title: "Convert Seurat objects"
output: html_notebook
---

Notebook for converting collected data into Seurat objects, when necessary. Datasets will not be saved within this project folder, only soft linked in data/expression



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(SeuratDisk)
library(qs)
```



# Tosches et al. 2018

Converting the Tosches et al. data into Seurat V3. See also [this thread on installing Seurat V2] (https://github.com/satijalab/seurat/issues/2625) for the V1 to V2 conversion (more notes in README.md).

Load Tosches V2 data

```{r}
pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Tosches2018/"
lizard_all = readRDS(paste0(pathToFiles, "v2obj/lizard.all.cells_v2.RDS"))
lizard_neurons = readRDS(paste0(pathToFiles, "v2obj/lizard.neurons.cells_v2.RDS"))

turtle_all = readRDS(paste0(pathToFiles, "v2obj/turtle.all.cells_v2.RDS"))
turtle_neurons = readRDS(paste0(pathToFiles, "v2obj/turtle.neurons.cells_v2.RDS"))
```

Update objects and save

```{r}
lizard_all_v3 = UpdateSeuratObject(lizard_all)
saveRDS(lizard_all_v3, file = paste0(pathToFiles, "v3obj/lizard_all_v3.RDS"))
lizard_neurons_v3 = UpdateSeuratObject(lizard_neurons)
saveRDS(lizard_neurons_v3, file = paste0(pathToFiles, "v3obj/lizard_neurons_v3.RDS"))

turtle_all_v3 = UpdateSeuratObject(turtle_all)
saveRDS(turtle_all_v3, file = paste0(pathToFiles, "v3obj/turtle_all_v3.RDS"))
turtle_neurons_v3 = UpdateSeuratObject(turtle_neurons)
saveRDS(turtle_neurons_v3, file = paste0(pathToFiles, "v3obj/turtle_neurons_v3.RDS"))
```



# La Manno et al. 2020
Load loom files

```{r}
p = "/links/groups/treutlein/USERS/tomasgomes/data/published/mousebrain/"
l5_all = Connect(filename = paste0(p, "l5_all.loom"), mode = "r")
dev_all = Connect(filename = paste0(p, "dev_all.loom"), mode = "r")

l5_all_seurat = as.Seurat(l5_all)
#dev_all_seurat = as.Seurat(dev_all)
```

Save objects

```{r}
saveRDS(l5_all_seurat, file = paste0(p, "l5_all_seurat.RDS"))
#saveRDS(dev_all_seurat, file = paste0(p, "dev_all_seurat.RDS"))
# dev data was ceded by Zhisong
```



# Raj et al. 2018

Similarly to Tosches data, this also has first to be converted from Seurat v1 to v2. I'm doing it in a different system with Seurat 2.3.4 installed.

Load Seurat v2 object

```{r}
p = "/links/groups/treutlein/USERS/tomasgomes/data/published/Raj2018_brain/"
fall_v2 = readRDS(paste0(p, "GSE105010_fall_v2.inDrops.RData"))
```

Update and save

```{r}
drerio_brain_v3 = UpdateSeuratObject(fall_v2)
saveRDS(drerio_brain_v3, file = paste0(p, "drerio_brain_v3.RDS"))
```



# Colquitt et al. 2021

```{r}
pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Colquitt2021/projects/HVC_RA_X/"
load(paste0(pathToFiles, "HVC_RA_X.RData"))
saveRDS(HVC_RA_X, file = paste0(pathToFiles, "HVC_RA_X.RDS"))

pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Colquitt2021/projects/X/"
load(paste0(pathToFiles, "MCX_COMBINED_C_seurat.RData"))
saveRDS(MCX_COMBINED_C_seurat, file = paste0(pathToFiles, "MCX_COMBINED_C_seurat.RDS"))
```



# Allen Brain Human

```{r}
pathToFiles = "/links/groups/treutlein/USERS/tomasgomes/data/published/Allen_human/"
dat10x = read.csv(paste0(pathToFiles, "matrix_10x.csv"), header = T, row.names = 1)
datSS = read.csv(paste0(pathToFiles, "matrix_SS.csv"), header = T, row.names = 1)
meta10x = read.csv(paste0(pathToFiles, "metadata_10x.csv"), header = T, row.names = 1)
metaSS = read.csv(paste0(pathToFiles, "metadata_SS.csv"), header = T, row.names = 1)
```

Make Seurat objects and save

```{r}
human_10x = CreateSeuratObject(counts = t(dat10x), meta.data = meta10x)
human_SS = CreateSeuratObject(counts = t(datSS), meta.data = metaSS)

saveRDS(human_10x, file = paste0(pathToFiles, "human_10x.RDS"))
saveRDS(human_SS, file = paste0(pathToFiles, "human_SS.RDS"))
```


