---
title: "Convert Seurat objects"
output: html_notebook
---

Notebook for converting collected data into Seurat objects, when necessary. Datasets will not be saved within this project folder, only soft linked in data/expression

Converting Tosches et al., 2018 data (and Seurat v1 in general):

Original data is Seurat V1
to convert to V3, first conversion to V2 is necessary
Thread on installing Seurat V2: satijalab/seurat#2625
use the UpdateSeuratObject function
then run it again for V2 to V3



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(SeuratDisk)
library(Matrix)
library(qs)
```



# Tosches et al. 2018

Converting the Tosches et al. data into Seurat V3. See also [this thread on installing Seurat V2] (https://github.com/satijalab/seurat/issues/2625) for the V1 to V2 conversion (more notes in README.md).

Load Tosches V2 data

```{r}
pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Tosches2018/"
lizard_all = readRDS(paste0(pathToFiles, "v2obj/lizard.all.cells_v2.RDS"))
lizard_neurons = readRDS(paste0(pathToFiles, "v2obj/lizard.neurons.cells_v2.RDS"))

turtle_all = readRDS(paste0(pathToFiles, "v2obj/turtle.all.cells_v2.RDS"))
turtle_neurons = readRDS(paste0(pathToFiles, "v2obj/turtle.neurons.cells_v2.RDS"))
```

Update objects and save

```{r}
lizard_all_v3 = UpdateSeuratObject(lizard_all)
saveRDS(lizard_all_v3, file = paste0(pathToFiles, "v3obj/lizard_all_v3.RDS"))
lizard_neurons_v3 = UpdateSeuratObject(lizard_neurons)
saveRDS(lizard_neurons_v3, file = paste0(pathToFiles, "v3obj/lizard_neurons_v3.RDS"))

turtle_all_v3 = UpdateSeuratObject(turtle_all)
saveRDS(turtle_all_v3, file = paste0(pathToFiles, "v3obj/turtle_all_v3.RDS"))
turtle_neurons_v3 = UpdateSeuratObject(turtle_neurons)
saveRDS(turtle_neurons_v3, file = paste0(pathToFiles, "v3obj/turtle_neurons_v3.RDS"))
```



# La Manno et al. 2020
Load file

```{r}
mousedev_filt = readRDS("/links/groups/treutlein/USERS/zhisong_he/Work/public_datasets/LaManno_mouse_dev_brain_atlas/dev_all_seurat.rds")
```

Filter and save

```{r}
mousedev_filt = mousedev_filt[,!(mousedev_filt$Class %in% c("Bad cells", "Undefined", "Blood", 
                                                            "Endoderm", "Ectoderm", "Mesoderm", 
                                                            "Fibroblast", "Gastrulation", 
                                                            "Mesenchyme", "Neural crest", 
                                                            "Neural tube", "Schwann cell")) &
                                !(mousedev_filt$ClusterName %in% c("nan")) &
                                mousedev_filt$nFeature_RNA>=250]

saveRDS(mousedev_filt, file = "/links/groups/treutlein/USERS/tomasgomes/data/published/LaManno2020_mouse/mousedev_filt.RDS")
```



# Zeisel et al., 2018
Load loom files

```{r}
p = "/links/groups/treutlein/USERS/tomasgomes/data/published/mousebrain/"
l5_all = Connect(filename = paste0(p, "l5_all.loom"), mode = "r")

l5_all_seurat = as.Seurat(l5_all)
```

Save objects

```{r}
saveRDS(l5_all_seurat, file = paste0(p, "l5_all_seurat.RDS"))
```


## Mouse dev Di Bella
Load data

```{r}
pp = "/links/groups/treutlein/USERS/tomasgomes/data/published/DiBella2021_mouse/"
clustertsne = read.table(paste0(pp, "cluster/cluster_scDevSC.merged.tsne.txt"), header = T)[-1,]
clusterumap = read.table(paste0(pp, "cluster/cluster_scDevSC.merged.umap.txt"), header = T)[-1,]
exp = Matrix::readMM(paste0(pp, "expression/601ae2f4771a5b0d72588bfb/gene_sorted-matrix.mtx.gz"))
bc = read.table(paste0(pp, "expression/601ae2f4771a5b0d72588bfb/barcodes.tsv"), header = F)
g = read.table(paste0(pp, "expression/601ae2f4771a5b0d72588bfb/genes.tsv"), header = F)
meta = read.table(paste0(pp, "metadata/metaData_scDevSC.txt"), header = T, sep = "\t", 
                  comment.char = "", quote = "")[-1,]
meta$Doublet_intersect[is.na(meta$Doublet_intersect)] = "Other"
rownames(meta) = meta$NAME
colnames(exp) = bc$V1
rownames(exp) = g$V1
srat = CreateSeuratObject(counts = exp, meta.data = meta)

srat = srat[,srat$scrublet_doublet=="FALSE" & meta$Doublet_intersect!="Doublet" &
              srat$New_cellType!="Low quality cells",]
saveRDS(srat, file = paste0(pp, "DiBella2021_mouseDev.RDS"))
```



# Raj et al. 2018

Similarly to Tosches data, this also has first to be converted from Seurat v1 to v2. I'm doing it in a different system with Seurat 2.3.4 installed.

Load Seurat v2 object

```{r}
p = "/links/groups/treutlein/USERS/tomasgomes/data/published/Raj2018_brain/"
fall_v2 = readRDS(paste0(p, "GSE105010_fall_v2.inDrops.RData"))
```

Update and save

```{r}
drerio_brain_v3 = UpdateSeuratObject(fall_v2)
# remove doublet clusters
drerio_brain_v3 = drerio_brain_v3[,!(drerio_brain_v3$Mod2.5 %in% c(24,31))]
saveRDS(drerio_brain_v3, file = paste0(p, "drerio_brain_v3.RDS"))
```



# Colquitt et al. 2021

```{r}
pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Colquitt2021/projects/HVC_RA_X/"
load(paste0(pathToFiles, "HVC_RA_X.RData"))
DefaultAssay(HVC_RA_X) = "RNA"
HVC_RA_X@assays$RNA@counts
saveRDS(HVC_RA_X, file = paste0(pathToFiles, "HVC_RA_X.RDS"))
finches_list = SplitObject(HVC_RA_X, split.by = "species")
saveRDS(finches_list$ZF, file = paste0(pathToFiles, "zfinch_seurat.RDS"))
saveRDS(finches_list$BF, file = paste0(pathToFiles, "bfinch_seurat.RDS"))

pathToFiles = "/links/groups/treutlein/USERS/Ashley/projects/axolotl/published/Colquitt2021/projects/X/"
load(paste0(pathToFiles, "MCX_COMBINED_C_seurat.RData"))
DefaultAssay(MCX_COMBINED_C_seurat) = "RNA"
saveRDS(MCX_COMBINED_C_seurat, file = paste0(pathToFiles, "MCX_COMBINED_C_seurat.RDS"))
```



# Allen Brain Human

```{r}
pathToFiles = "/links/groups/treutlein/USERS/tomasgomes/data/published/Allen_human/"
dat10x = read.csv(paste0(pathToFiles, "matrix_10x.csv"), header = T, row.names = 1)
datSS = read.csv(paste0(pathToFiles, "matrix_SS.csv"), header = T, row.names = 1)
meta10x = read.csv(paste0(pathToFiles, "metadata_10x.csv"), header = T, row.names = 1)
metaSS = read.csv(paste0(pathToFiles, "metadata_SS.csv"), header = T, row.names = 1)
```

Make Seurat objects and save

```{r}
human_10x = CreateSeuratObject(counts = t(dat10x), meta.data = meta10x)
human_SS = CreateSeuratObject(counts = t(datSS), meta.data = metaSS)

saveRDS(human_10x, file = paste0(pathToFiles, "human_10x.RDS"))
saveRDS(human_SS, file = paste0(pathToFiles, "human_SS.RDS"))
```



# Cebrian-Silla et al., 2021
Load files

```{r}
pathToFiles = "/links/groups/treutlein/USERS/tomasgomes/data/published/CebrianSilla2021_mouse/"
nuc_dat = readRDS(paste0(pathToFiles, "sNucSeq.rds"))
sc_dat = readRDS(paste0(pathToFiles, "scSeq.rds"))
nuc_meta = read.csv(paste0(pathToFiles, "meta_nuc.tsv"), header = T, sep = "\t", row.names = 1)
sc_meta = read.csv(paste0(pathToFiles, "meta_sc.tsv"), header = T, sep = "\t", row.names = 1)
```

Add metadata, filter, and save

```{r}
sc_dat@meta.data = sc_meta
nuc_dat@meta.data = nuc_meta

saveRDS(nuc_dat, file = paste0(pathToFiles, "nuc_dat.RDS"))
saveRDS(sc_dat, file = paste0(pathToFiles, "sc_dat.RDS"))
```









